---
title: "Create a Dynamic Timeline using RStudio"
author: "Lok Yiu"
date: "20 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create a Timeline in R
This is an example of creating a timeline using the R script created by Dean Attali. [Link](https://daattali.com/shiny/timevis-demo/) and the `timevis` package. I really like it as the script is very simple ad tidy. This is a great example for data scientists or analysts using our programming knowledge to make thing in the way we want (and without any additional cost to purchase other software)! 

My approach is slightly different to Dean's one. I think making a task list on MS Excel is more handy than typing the task and dates directly into the script, in particular I know my list will grow dramatically in the future. Thus I have made a modification in this demo: Making a timeline by reading the a task list from a CSV file. 

To begin with the very first step is to install RStudio [Link](https://www.rstudio.com/), then install and load the libraries.

```{r library, warning=FALSE,message=FALSE}
# install.packages("shiny") # Remove the "#" in this line and run if you have not installed the packages
# install.packages("timevis") # Remove the "#" in this line and run if you have not installed the packages
library("shiny")
library("timevis")
```

## Load Data
As an example, you can download my example project task lists on Google drive:

```{r google}
id <- "1N7Nyi-u-mqAkQ8SmwsgK6D4QkGE0kOpK"
dt <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id))
print(dt)
```

## Prepare the task list
I reformat 3 columns to make sure the date and task can displayed properly.

```{r prepare}
dt[dt==""] <- NA # Replace all blank to NA
dt$start <- as.Date(as.character(dt$start), "%d/%m/%Y") # reformat the date field to my preferred format
dt$end <- as.Date(as.character(dt$end), "%d/%m/%Y")
dt$content <- as.character(dt$content) # formate the content as text characters
```

## Plot the Timeline
Here is the exciting part. Simply run the following code and we will have a timeline. You can zoom in and out and drag horizontally to display the tasks. 

```{r timeline1}
timevis(dt, width = 900)
```

## Plot the Timeline with grouping
Here I made a small advancement by visualising the tasks by groups. I hope you enjoy this demostration.

```{r timeline2}
groups = data.frame(id = 1:4, content = c("Outreaching", "System/Platform Changes", "Xmas Shutdown Period", "Key Milestone"))
timevis(dt, groups, width = 900)
```

